\documentclass[11pt]{article}
\usepackage[margin = 1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{enumitem}
\usepackage{url}
\usepackage[parfill]{parskip}
\newcommand{\skipline}{\vspace{\baselineskip}}
\newenvironment{problem}[1]{\textbf{Problem #1: }}{\newpage}


\begin{document}
	
	\begin{center}
		\textbf{Quiz 1} \\
		\textbf{Ordinary Differential Equations} \\
		\textbf{Math 537} \\
		\textbf{Stephen Giang} \\
	\end{center}
	\skipline
	\begin{problem}{1}
		Consider the following first-order ODE's: 
		Please provide one example for each of the below ODEs and discuss the
		corresponding solutions.
		\begin{enumerate}[label = (\alph*)]
			\item Separable ODEs
			\begin{align*}
				\frac{dy}{dt} &= yt & \ln(y) &= \frac{t^2}{2} + C\\
				\frac{dy}{y} &= t\,dt & y &= Ce^{\frac{t^2}{2}}
			\end{align*}
			For Separable ODEs, they will be in the form of $\frac{dy}{dt} = f(t,y) = p(t)q(y)$, with solutions 
			$$ y = g(t) $$
			\item Linear ODEs
			\begin{align*}
				\frac{dy}{dt} + y &= e^t & \frac{d}{dt}(e^{t} y) &= e^{2t} & y = \frac{1}{2}e^t + Ce^{-t}\\
				e^t \frac{dy}{dt} + e^t y &= e^{2t} & e^{t} y &= \frac{1}{2}e^{2t} + C
			\end{align*}
			For Linear ODEs, they will be in the form of $\frac{dy}{dt} + p(t)y = g(t)$, with solutions of 
			$$y(t) = e^{-\int p(t)\,dt} \left[\int e^{\int p(t)\,dt}g(t)\,dt + C\right]$$
			\item Exact ODEs
			\begin{align*}
				2xy^2 + 2yx^2\frac{dy}{dt} = 0 
			\end{align*}
			Notice that it is in the form of $M(x,y) + N(x,y)y' = 0$, with $M_y(x,y) = 4xy = N_x(x,y)$, thus giving us the solution ($\phi(x,y)$):
			\begin{align*}
				\int 2xy^2 \, dx &= x^2y^2 + h(y) \\
				\int 2yx^2 \, dy &= x^2y^2 + g(x) \\
				\phi(x,y) &= x^2y^2 + C
			\end{align*}
			For Exact ODEs, they will be in the form of $M(x,y) + N(x,y)y' = 0$, with $M_y(x,y) = N_x(x,y)$, with solutions of
			$$\phi(x,y)=\int M(x,y)\,dx = \int N(x,y)\,dy$$ 
			\newpage
			\item Bernoulli Equations
			\begin{align*}
				\frac{dy}{dt} + y = y^2
			\end{align*}
			Notice we can substitute $u=y^{1-2} = \frac{1}{y}$, $\frac{du}{dt} = \frac{-1}{y^2}\frac{dy}{dt}$ into the equation after multiplying the equation by $\frac{-1}{y^2}$:
			\begin{align*}
				\frac{du}{dt} - u &= -1 & \ln(u-1) &= t + C \\
				\frac{du}{u - 1} &= dt & u &= Ce^{t} + 1
			\end{align*}
			Now we need to resubstitute $u=y^{1-2}$, and we get:
			\begin{align*}
				\frac{1}{y} &= Ce^{t} + 1 \\
				y &= \frac{1}{Ce^{t} + 1}
			\end{align*} 
		\end{enumerate}
	\end{problem}

	\begin{problem}{2}
		Consider the following homogeneous linear 2nd-order ODEs
		with constant coefficients:
		$$y'' + ay' + by = 0$$
		where a and b are constant. Please discuss three types of solutions based on
		the so-called characteristic equation. \\

		By converting this equation into a system of first order differential equations.  We can then find the characteristic equation to be $\lambda^2 + a\lambda + b = 0$.  We can find the roots $\lambda_1$ and $\lambda_2$ from the characteristic equations. Notice the three types of solutions: \\
			\begin{enumerate}[label = (\alph*)]
				\item Distinct Real Roots with ($\lambda_1 \not = \lambda_2$):
				$$y(t) = c_1e^{\lambda_1 t} + c_2e^{\lambda_2 t}$$
				\item Equal Real Roots with ($\lambda_1 = \lambda_2$):
				$$y(t) = c_1e^{\lambda_1 t} + c_2te^{\lambda_2 t}$$
				\item Roots with ($\lambda_{1,2} = \mu \pm i\nu$):
				$$c_1e^{\mu t}\cos(\nu t) + c_2e^{\mu t}\sin(\nu t)$$
			\end{enumerate}
	\end{problem}

	\begin{problem}{3}
		Consider the following Euler-Cauchy equation:
		$$x^2y'' + axy' + by = 0 $$
		where a and b are constant.
		\begin{enumerate}[label = (\alph*)]
			\item  Please discuss three types of solutions:
			We can substitute $y(x) = x^r$.  This will get us its characteristic equation to be $r^2 + (a-1)r + b$ and get 2 solutions $r_1$ and $r_2$ with $x > 0$:
				\begin{enumerate}[label = (\roman*)]
					\item Distinct Real Roots with ($r_1 \not = r_2$):
					$$y(t) = c_1x^{r_1} + c_2x^{r_2}$$
					\item Equal Real Roots with ($r_1 = r_2$):
					$$y(t) = (c_1 + c_2\ln x)x^{r_1}$$
					\item Roots with ($r_{1,2} = \mu \pm i\nu$):
					$$y(t) = x^\mu \left[c_1 \cos(\nu \ln x) + c_2 \sin (\nu \ln x)\right]$$
				\end{enumerate} 
			\item Introduce a new independent variable $(t), x = e^t$, to convert the above 
			Euler-Cauchy equation into a second-order ODE with constant coefficients (i.e., in the form of Eq. 2).
			\\ \\
			Notice the following with $x = e^t$:
				\begin{align*}
					y' &= \frac{dy}{dx} = \frac{\frac{dy}{dt}}{\frac{dx}{dt}} = \frac{dy}{dt}e^{-t} \\
					y'' &= \frac{d^2y}{dx^2} = \left(\frac{d}{dx}\right) \frac{dy}{dx} = \frac{\frac{d}{dt}}{ \frac{dx}{dt}}\left(\frac{dy}{dx}\right) \\
					&= \frac{\frac{d}{dt} \frac{dy}{dx}}{\frac{dx}{dt}} = \frac{e^{-t}\frac{d^2y}{dt^2} - e^{-t}\frac{dy}{dt}}{e^t} \\
					&= e^{-2t}\left(\frac{d^2y}{dt^2} - \frac{dy}{dt}\right)
				\end{align*}
			Now we can resubsitute the following into our Cauchy-Euler equation with $x = e^t$:
				\begin{align*}
					&e^{2t}\left(e^{-2t}\left(\frac{d^2y}{dt^2} - \frac{dy}{dt}\right)\right) + ae^t\left(\frac{dy}{dt}e^{-t}\right) + by = 0 \\
					&\frac{d^2y}{dt^2} + (a-1)\frac{dy}{dt} + by = 0
				\end{align*}
			This looks very similar to our characteristic equation as noticed in (a).
		\end{enumerate}
	\end{problem}

	\begin{problem}{4}
		Provide a brief summary on what has been completed in this
		assignment.
		\\ \\
		In this assignment, we reviewed the techniques to solve first-order ODEs, the different type of solutions from 2nd-order ODE's, the different types of solutions from Cauchy-Euler Equations and how the Cauchy-Euler Equation can be converted into a constant-coefficient 2nd Order ODE.
	\end{problem}


\end{document}
