\documentclass[12pt]{article}
\usepackage[margin = 1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{cancel}

\begin{document}
	
	\begin{center}
		\textbf{Homework 3.2} \\
		\textbf{Linear Algebra} \\
		\textbf{Math 524} \\
		\textbf{Stephen Giang} \\
	\end{center}

\noindent \textbf{Section 3.D Problem 2: } Suppose V is finite-dimensional and dim $V > 1$. Prove that the set of
noninvertible operators on V is not a subspace of $\mathcal{L} (V)$.

	\begin{proof}[Solution 3.D.2]
		Let V be finite-dimensional with dim $V = n$, \quad $n \in \mathbb{Z}^{> 1}$. \\
		Let \{$v_1, ..., v_n$\} be a basis of V. Let $S,T \in \mathcal{L}(V)$
		\begin{align*}
			Sv_1 = v_1 &&& Sv_2 = 0 && Sv_k = 0 \\
			Tv_1 = 0 &&& Tv_2 = v_2 && Tv_k = v_k 
		\end{align*}
		So $S,T$ are both noninvertible operators on V, such that S is not injective, and T is not surjective
		\begin{align*}
			&(S+T)v_1 = Sv_1 + Tv_1 = v_1 \\
			&(S+T)v_k = Sv_k + Tv_k = v_k 
		\end{align*}
		Because (S+T) is injective and surjective, (S+T) is invertible, so S,T is not closed under addition, thus they are not a subspace of $\mathcal{L}(V)$ \\
	\end{proof}

\vspace{\baselineskip}
\vspace{\baselineskip}

\noindent \textbf{Section 3.D Problem 3: } Suppose V is finite-dimensional, U is a subspace of V, and $S \in \mathcal{L}(V)$. Prove there exists an invertible operator $T \in \mathcal{L}(V)$ such that $Tu = Su$ for every $u \in U$ if and only if S is injective.

	\begin{proof}[Solution 3.D.3 ]
		Let V be finite-dimensional, $U \subseteq V$, and $S \in \mathcal{L}(V)$.
		\\
		($=>$) Let $T \in \mathcal{L}(V)$ be invertible, and $Tu = Su$, $\forall u \in U$. 
		\\ \\
		Because T is invertible, T is injective, such that Tu = 0, u is exclusively 0. And because Tu = Su = 0 only when u is exclusively 0, S is injective as well.	
		\\ \\
		($<=$) Let S be injective, and \{$u_1,...,u_m,v_1,...,v_n$\} be an extended Basis of V from U. 
		\\ \\
		Because S is injective, we can have \{$Su_1,...,Su_m,w_1,...,w_n$\} be an extended Basis of V from S. Let T be defined as: 
		\begin{align*}
			&Tu_k = Su_k \quad 1 \leq k \leq m \\
			&Tv_j = w_j  \quad 1 \leq j \leq n
		\end{align*}
		Because T maps the entire basis \{$u_1,...,u_m,v_1,...,v_n$\} to another basis entirely, \{$Su_1,...,Su_m,w_1,...,w_n$\}, T is invertible $\in \mathcal{L}(V)$. \\
	\end{proof}

\vspace{\baselineskip}
\vspace{\baselineskip}

\noindent \textbf{Section 3.E Problem 2: } Suppose $V_1, ..., V_m$ are vector spaces such that $V_1 \times \dots \times V_m$ is finite dimensional. Prove that $V_j$ is finite-dimensional for each $j = 1, ..., m $

	\begin{proof}[Solution 3.E.2]
		Let  $V_1, ..., V_m$ be vector spaces such that $V_1 \times \dots \times V_m$ is finite dimensional.
		\\ \\
		By Theorem 3.76, Dim ($V_1 \times \dots \times V_m$) = $\sum_{k = 1}^{m}$Dim ($V_k$). 
		\\ \\
		If any $V_k$ was not finite dimensional, then the dimension of $V_k$ would be infinite, such that the sum with the other dimensions would not be finite, as integers are closed under addition. \\
	\end{proof}

\vspace{\baselineskip}
\vspace{\baselineskip}

\noindent \textbf{Section 3.E Problem 4: } Suppose $V_1, ..., V_m$ are vector spaces.\\Prove that $\mathcal{L} (V_1 \times \dots \times V_m, W)  $ and $\mathcal{L}(V_1, W) \times \dots \times \mathcal{L} (V_m,W) $ are isomorphic vector spaces.
	
	\begin{proof}[Solution 3.E.4]
		Let $V_1, ..., V_m$ be vector spaces. 
		\\ \\
		Let $f \in \mathcal{L} (V_1 \times \dots \times V_m, W)$, with $ f_i: V_i \rightarrow W $, $$f_i(v_i) = f(0,0,0,v_i,0,0,0) \in \mathcal{L}(V_i,W) \quad 1 \leq i \leq m$$
		
		Let $\Gamma$ : $\mathcal{L} (V_1 \times \dots \times V_m, W) \rightarrow \mathcal{L}(V_1, W) \times \dots \times \mathcal{L} (V_m,W)$ such that $$\Gamma(f) = (f_1 , ..., f_m)$$
		
		Let $\Gamma^{-1}$ : $\mathcal{L}(V_1, W) \times \dots \times \mathcal{L} (V_m,W) \rightarrow \mathcal{L} (V_1 \times \dots \times V_m, W)$ such that $$\Gamma^{-1} (f_1, ..., f_m) = (f_1(v_1), ... , f_m(v_m))$$
		
		Because $\Gamma$ is linear and has an inverse, $\mathcal{L} (V_1 \times \dots \times V_m, W)  $ and $\mathcal{L}(V_1, W) \times \dots \times \mathcal{L} (V_m,W) $ are isomorphic vector spaces. \\
	\end{proof}
\end{document}
